Applied Measure Theory for Probabilistic Modeling

### 2022-05-12 paper/paper.tex
edits

\section{Base measures and density recursion}
It's very useful for users to be able to \emph{call} \verb|logdensity_rel(μ, ν, x)|. But \emph{defining} things directly in these terms would require a definition for every pair of possible measures, which is of course intractable.

@@

A Markov kernel, in particular, corresponds to the distribution of a \emph{parametrized random variable}, a function with random outcomes (or to ``mechanisms'' in causal inference, for example in \cite{https://doi.org/10.5281/zenodo.1005091}). For example in \verb|MeasureTheory|, the kernel
\[
\kappa\colon x \mapsto \text{Normal}(x,\sqrt{x})
\]
can be expressed as
\begin{verbatim}
    κ = kernel(Normal) do x
        (μ=x, σ=√x)
    end
\end{verbatim}
or equivalently
\begin{verbatim}
    κ = kernel(Normal; μ=identity, σ=sqrt)
\end{verbatim}
corresponding to the random function
\begin{verbatim}
    f(rng, x) = x + √x * randn(rng)
\end{verbatim}

In this example, working in terms of a \verb|kernel| allows us to make the linear relationship between the argument \verb|x| and \verb|mean(κ(x))| transparent.

@@

The above \verb|Likelihood| type is specified using a kernel and observed data. Though it is not an \verb|AbstractMeasure|, we include a \verb|logdensity| method for computational convenience. The core of the implementation is then
\begin{verbatim}
    struct Likelihood{F,S,X}
        k::Kernel{F,S}
        x::X
    end
\end{verbatim}

>>>>

\section{Base measures and density recursion}
It's very useful for users to be able to \emph{call} \verb|logdensity_rel| for arbitrary measures. But \emph{defining} things directly in these terms would require a definition for every pair of possible measures, which is of course intractable.

@@

A Markov kernel, in particular, corresponds to the distribution of a \emph{parametrized random variable}, a function with random outcomes (or to ``mechanisms'' in causal inference, for example in \cite{https://doi.org/10.5281/zenodo.1005091}). For example, the kernel
\[
\kappa\colon x \mapsto \text{Normal}(x,\sqrt{x})
\]
corresponds to the random function
\begin{verbatim}
    f(rng, x) = x + √x * randn(rng)
\end{verbatim}
and can be expressed in \verb|MeasureTheory| as
\begin{verbatim}
    κ = kernel(Normal) do x
        (μ=x, σ=√x)
    end
\end{verbatim}
or
\begin{verbatim}
    κ = kernel(Normal; μ=identity, σ=sqrt)
\end{verbatim}

The latter formulation decomposes the kernel into separate ``parameter maps'', making relationships like the one between \verb|x| and \verb|mean(κ(x))| more explicit.

@@

The above \verb|Likelihood| type is specified using a kernel and observed data. Though it is not an \verb|AbstractMeasure|, we include a \verb|logdensity| method for computational convenience. The core of the implementation is then
\begin{verbatim}
    struct Likelihood{K,X}
        k::K
        x::X
    end
\end{verbatim}
